---
title: "Lab 13"
author: "Andrew Graves"
date: "April 23/24 2020"
output:
  ioslides_presentation:
    smaller: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.height = 3.5)
```

# Specification of random-effects structures with lmer

## Load the data

```{r load}
library(tidyverse)
library(haven) # For reading SPSS files
library(lmerTest) 
# lmerTest is a direct wrapper for lme4; it provides approximate df
# which often can help diagnose model specifiation problems
options(contrasts = c("contr.sum","contr.poly"))

# Set a reasonable optimizer
my_opt <- lmerControl(optimizer = "bobyqa",
                      optCtrl = list(maxfun = 2e6))

# The popularity dataset contains characteristics of pupils in different classes
# Borrowed from this blogpost: https://www.rensvandeschoot.com/tutorials/lme4/
dat <- read_sav(file = 
  paste0("https://github.com/MultiLevelAnalysis",
         "/Datasets-third-edition-Multilevel-book/blob/master",
         "/chapter%202/popularity/SPSS/popular2.sav?raw=true")
  ) %>%
  select(pupil:popular) %>%
  mutate_at(vars(extrav, texp, popular), scale) %>%
  mutate_at(vars(pupil, class), factor) %>%
  mutate(sex = factor(case_when(
    sex == 0 ~ "boy",
    TRUE ~ "girl"
  ))) %>%
  mutate_if(is.numeric, as.vector)
```

## Explore the data structure

```{r str}
glimpse(dat)
```

## One-way RM ANOVA style

```{r mod0}
mod0 <- lmer(popular ~ extrav + (1|class/extrav), 
            data = dat, REML = FALSE, control = my_opt)
summary(mod0)$coefficients
```

## Lmer style with random slopes (analagous to latent-growth curve)

```{r mod1}
mod1 <- update(mod0, formula = popular ~ extrav + (extrav|class))
summary(mod1)$coefficients
head(ranef(mod1)[[1]])
```

## Remove the intercept and just keep the slope

```{r mod2}
mod2 <- update(mod0, formula = popular ~ extrav + (0 + extrav|class))
summary(mod2)$coefficients
head(ranef(mod2)[[1]])
```

## Remove the correlation between the slope and intercept

```{r mod3}
mod3 <- update(mod0, formula = popular ~ extrav + (extrav||class))
summary(mod3)$coefficients
head(ranef(mod3)[[1]])
```

## Model comparison

```{r mod_comp}
anova(mod0, mod1, mod2, mod3)
```

## We can fit various types of models

```{r vary}
mod4 <- update(mod0, formula = popular ~ extrav + (extrav + sex|class))

set.seed(42)
mod5 <- lmer(formula = popular ~ extrav + sex + (extrav + sex|class),
             data = dat, REML = FALSE,
             lmerControl(optimizer = "nmkbw",
                         optCtrl = list(maxfun = 2e6)))
set.seed(42)
mod6 <- update(mod5, formula = popular ~ extrav + sex + texp + 
              (extrav + sex|class))
```

# Now it's your turn!

Download the in-class activity and answer the questions as best you can. 

Collaborate with your peers to write code. 

I am available for programming questions.