---
title: "PSYC 7720 Lab" 
subtitle: "Lab 10 Activity"
author: "Andrew Graves, Department of Psychology, University of Virginia"
output:
  pdf_document: default
---

## Directions:

A.	Download the angle_noise_data.csv. This is a two-factor within subjects design
with two levels of the noise factor and three levels of the angle factor. In the data, abs represents the condition where noise is absent and pres represents where noise is present. The 0, 4, and 8 denote the three different angle conditions. The measurements in each cell are aggregated reaction times.
B.	Answer the following questions and save the code you used in an R script.
C.	You have until the end of lab to complete.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, fig.height = 4)
```

```{r init}
library(tidyverse)
library(ez)
library(emmeans)
options(contrasts = c("contr.sum","contr.poly"))
theme_set(theme_classic())
```

## Questions:

1. Convert the angle_noise_data from wide format to long format. Name the first column *id*, the second column *noise*, the third column *angle*, the fourth column *rt*, and the dataframe *long_data*. Print the head and the dimensions of the long data.

```{r 1}
long_data <- read.csv("angle_noise_data.csv") %>%
  rownames_to_column(var = "id") %>%
  pivot_longer(-id,
               names_to = "ws", values_to = "rt") %>%
  separate(ws, into = c("noise", "angle"), sep = -1,
           remove = TRUE) %>%
  mutate_if(is_character, factor) %>%
  mutate(angle = factor(angle, ordered = TRUE))
head(long_data)
dim(long_data)
```

2. Ensure that the within-subjects conditions are appropriately implemented as qualitative and/or quantitative factors. Print the appropriate structure of *long_data* using str().

```{r 2}
str(long_data)
```

3. Run and interpret the results of the two-way factorial RM ANOVA using aov.

```{r 3}
aov_mod <- aov(rt ~ noise * angle + Error(id/(noise * angle)), data = long_data)
summary(aov_mod)
```

4. Run and interpret the results of the two-way factorial RM ANOVA using ezANOVA.

```{r 4}
 ez_mod <- ezANOVA(data = long_data,
                  dv = rt,
                  within = .(angle, noise),
                  wid = id,
                  type = 3,
                  return_aov = TRUE,
                  detailed = TRUE
        )
ez_mod

# Check the residuals for normality
resid <- proj(ez_mod$aov)[[3]][, "Residuals"]

resid %>%
  tibble() %>%
  ggplot(aes(sample = `.`)) + 
  geom_qq() + 
  stat_qq_line() +
  labs(x = "Theoretical Quantiles",
       y = "Sample Quantiles")

# Interaction is significant... here we are 
# testing the effect of noise within each 
# level of angle. You could also test the
# effect of noise at each angle as a 
# follow-up for this interaction.
follow_up <- contrast(
  emmeans(ez_mod$aov, ~ noise|angle),
          method = "trt.vs.ctrl",
          adjust = "bonferroni"
  )
follow_up

ez_plot <- ezPlot(
  data = long_data,
  dv = rt,
  wid = id,
  within = .(angle, noise),
  x = angle,
  split = noise,
  x_lab = "Noise",
  split_lab = "Angle",
  y_lab = "RT"
  )
ez_plot
```

- There was a significant two-way interaction between *noise* and *angle* on *rt*. Following up this interaction, it seems that the contrasts between the two noise conditions is not significant when *angle* is 0, but is significant when *angle* is 4 and 8. Another way to interpet this interaction is that there is a stronger positive linear trend when *noise* is present relative to when it is absent.