---
title: "Lab 10"
author: "Andrew Graves"
date: "April 2/3 2020"
output:
  ioslides_presentation:
    smaller: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.height = 3.5)
```

# Introduction to Factorial RM ANOVA in R

## Load the data

```{r load}
library(tidyverse)
library(ez)
library(lme4)
library(effects)

# Making sure we have ANOVA-appropriate contrasts:
options(contrasts = c("contr.sum","contr.poly"))

# Data from the paper "Modeling the dynamics of recognition memory testing 
# with a combined model of retrieval and decision making
# doi: 10.17632/226zz69bk9.1
# ----------
# Manipulates list length between 24 and 96 items, 
# normative word frequency (LF vs. HF), 
# and the length of the retention interval.

dat <- read_csv("exp1testing.csv") %>%
  mutate_if(is.character, factor) %>%
  mutate(subj = factor(subj),
         RT = log(RT))
```

## Explore the data structure

```{r str}
glimpse(dat)
```

## Compute subject * condition aggregate means for Factorial RM ANOVA

```{r agg}
grp_dat <- dat %>%
  group_by(subj, wf, length) %>%
  summarize(mean_rt = mean(RT))

glimpse(grp_dat)
```

# Factorial RM ANOVA on word frequency and list length

## Use aov and Error for Factorial RM ANOVA

```{r aov_fact}
# Base r solution
aov_fact <- aov(mean_rt ~ wf * length + 
               Error(subj/(wf * length)), data = grp_dat)
```

## View aov solution

```{r aov_fact_soln}
summary(aov_fact)
```

## Use ezANOVA for Factorial RM ANOVA

```{r ez_fact}
# ez package solution (checks for violations of sphericity)
# ----------
ez_fact <- ezANOVA(data = grp_dat,
                  dv = .(mean_rt),
                  within = .(wf, length),
                  wid = .(subj),
                  detailed = TRUE
        )
ez_fact
```

## Use lmer (on aggregate data) for Factorial RM ANOVA

```{r lmer_agg}
set.seed(42)
grp_lmer <- lmer(mean_rt ~ wf * length + 
                (1|subj/wf) + (1|subj/length) + (1|subj/wf:length),
                data = grp_dat,
                control=lmerControl(check.nobs.vs.nRE="ignore",
                check.nobs.vs.nlev = "ignore",
                optimizer = "Nelder_Mead",
                optCtrl = list(maxfun = 1e5)
                   ))
```

## View the lmer aggregate solution

```{r lmer_agg_soln}
summary(grp_lmer)$coefficients
anova(grp_lmer)
```

# Model the full data (trial-level) with lmer

## Factorial RM ANOVA on trial-level data

```{r lmer_fact}
set.seed(42)
full_lmer_fact <- lmer(RT ~ wf * length + (1|subj/wf) + (1|subj/length) +
                  (1|subj/wf:length), data = dat)

fit_all_fact <- allFit(full_lmer_fact)
diff_opts_fact <- summary(fit_all_fact)

set.seed(42)
final_fact <- update(full_lmer_fact, 
                     control = lmerControl(optimizer = "bobyqa",
                                           optCtrl = list(maxfun = 2e6)))
```

## Messages from the various optimizer solutions

```{r optim1}
diff_opts_fact$msgs
```

## View fixed effects output

```{r lmer_fact_res}
summary(final_fact)$coefficients
anova(final_fact)
```

## Plot the factorial lmer model

```{r effects1}
# allEffects is a function within the effects package: Extracts model parameters
plot(allEffects(final_fact))
```

# Adding a between-subjects factor

## Create an age factor (old = 0, young = 1)

```{r age}
subj <- grp_dat %>%
  arrange(desc(mean_rt)) %>%
  ungroup() %>%
  select(subj) %>%
  unique() %>%
  filter(subj != 0)

is_young <- rep(c(0, 1), each = nrow(subj) / 2) %>%
  factor()
joiner <- cbind(subj, is_young)
head(joiner)

age_dat <- dat %>%
  inner_join(joiner, by = "subj")
```

## Fit the between-subjects factor (additive) in lmer

```{r lmer_fact_bet}
set.seed(42)
full_lmer_fact_bet <- lmer(RT ~ wf * length + is_young + 
                          (1|subj/wf) + (1|subj/length) + (1|subj/wf:length),
                          data = age_dat)

set.seed(42)
fit_all_fact_bet <- allFit(full_lmer_fact_bet)
diff_opts_fact_bet <- summary(fit_all_fact_bet)
diff_opts_fact_bet$msgs

set.seed(42)
final_fact_bet <- update(full_lmer_fact_bet, 
                     control = lmerControl(optimizer = "bobyqa",
                                           optCtrl = list(maxfun = 2e6)))
```

## Model now incorporates age

```{r effects2}
plot(allEffects(final_fact_bet))
```

## Add between subjects factor (interactive) to aov

```{r aov_fact_bet}
age_grp_dat <- grp_dat %>%
  inner_join(joiner, by = "subj") %>%
  ungroup() %>%
  mutate(subj = factor(subj))

aov_fact_bet <- aov(mean_rt ~ wf * length * is_young + 
               Error(subj/(wf * length)), data = age_grp_dat)
```

## Add between subjects factor (interactive) to ezANOVA

```{r ez_fact_bet}
ez_fact_bet <- ezANOVA(data = age_grp_dat,
                  dv = mean_rt,
                  within = .(wf, length),
                  wid = subj,
                  between = is_young
)
ez_fact_bet
```

# Now it's your turn!

Download the in-class activity and answer the questions as best you can. 

Collaborate with your peers to write code. 

I am available for programming questions.