---
title: "Lab 12"
author: "Andrew Graves"
date: "April 16/17 2020"
output:
  ioslides_presentation:
    smaller: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.height = 3.5)
```

# Introduction to Multivariate approach to RM ANOVA in R

## Load the data

```{r load}
library(tidyverse)
library(car)

# Making sure we have ANOVA-appropriate contrasts:
options(contrasts = c("contr.sum","contr.poly"))

# Hypothetical data regarding the role of 
# premature births on the growth of newborn
# infants. Infants were weighed at the same 
# time of day for 4 consecutive days.
baby_dat <- read_csv("baby_weights.csv") %>%
  mutate(Premature, factor(Premature))

# Same angle and noise data from Lab 10 
# activity, but with the added between-subject 
# factor of doing the task either in a dark or 
# bright room
rt_dat <- read_csv("angle_noise_lighting_data.csv") %>%
  mutate(lighting = factor(lighting))
```

## Explore the data structure

```{r str}
glimpse(baby_dat)
glimpse(rt_dat)
```

# One-way multivariate RM ANOVA

## Dependent-variable matrix for the one-way model

```{r build_mod1a}
# Dependent-variable matrix
dvs_1 <- baby_dat %>%
  select(Day1:Day4)

# The model
lm(as.matrix(dvs_1) ~ 1)
```

## The names of the within-subjects factors as well as the design for the one-way model
```{r build_mod1b}
# The names of the factors
data.frame(days = names(dvs_1))

# The RM design
~days
```

## Use car's Anova function to build the one-way model

```{r run_mod1}
mod1 <- car::Anova(mod = 
           lm(as.matrix(dvs_1) ~ 1),
           idata = data.frame(days =
           names(dvs_1)),
           idesign = ~days,
           type = 3
           )
summ_mod1 <- summary(mod1)
```

## View car's Anova solution for days effect

```{r mod1_soln}
summ_mod1$multivariate.tests$days
```

# Two-way multivariate RM ANOVA

## Dependent-variable matrix for the two-way model

```{r build_mod2a}
# Dependent-variable matrix
dvs_2 <- rt_dat %>%
  select(abs0:pres8)

# The model
lm(as.matrix(dvs_2) ~ 1)
```

## The names of the within-subjects factors as well as the design for the two-way model

```{r build_mod2b}
# The names of the factors; multiple ways to 
# do this!
dv_names <- str_split(names(dvs_2), "s",
                      simplify = TRUE)

colnames(dv_names) <- c("noise", "angle")

# The RM design
~noise * angle
```

## Use car's Anova function to build the two-way model

```{r run_mod2}
mod2 <- car::Anova(mod = lm(as.matrix(dvs_2) ~ 1),
                  idata = data.frame(dv_names),
                  idesign = ~ noise * angle,
                  type = 3
)
summ_mod2 <- summary(mod2)
```

## View car's Anova solution for noise effect

```{r mod2_soln1}
summ_mod2$multivariate.tests$noise
```

## Now angle...

```{r mod2_soln2}
summ_mod2$multivariate.tests$angle
```

## Now noise x angle interaction...

```{r mod2_soln3}
summ_mod2$multivariate.tests$`noise:angle`
```

# Adding a between-subjects factor for multivariate RM ANOVA

## Build the full interaction model

```{r build_mod3}
mod3 <- car::Anova(
  mod = 
    lm(cbind(
      abs0, abs4, abs8, pres0, pres4, pres8) ~ 
        lighting, data = rt_dat),
        idata = data.frame(dv_names),
        idesign = ~ noise * angle, 
        type = 3
)
summ_mod3 <- summary(mod3)
```

## View car's Anova solution for three-way interaction

```{r mod3_soln}
summ_mod3$multivariate.tests$`lighting:noise:angle`
```

# Now it's your turn!

Download the in-class activity and answer the questions as best you can. 

Collaborate with your peers to write code. 

I am available for programming questions.