---
title: "Lab 9"
author: "Andrew Graves"
date: "March 26/27 2020"
output:
  ioslides_presentation:
    smaller: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.height = 3.5)
```

# Introduction to RM ANOVA in R

## The sleepstudy dataset

```{r sleepstudy}
library(tidyverse)
library(ez)
library(lme4)
theme_set(theme_classic())
theme_update(text = element_text(family = "serif"))

# Making sure we have ANOVA-appropriate contrasts:
options(contrasts = c("contr.sum","contr.poly"))

# The average reaction time per day for subjects in a sleep deprivation study. 
# On day 0 the subjects had their normal amount of sleep. Starting that night 
# they were restricted to 3 hours of sleep per night. The observations 
# represent the average reaction time on a series of tests given each 
# day to each subject.

# Run ?sleepstudy for more information
data(sleepstudy)
str(sleepstudy)
```

## Create a time factor (Period) at lower resolution for demonstration purposes

```{r period}
period <- rep(c(1:5), 
              times = nrow(sleepstudy) / length(unique(sleepstudy$Days)), 
              each = length(unique(sleepstudy$Days)) / 5
              )

dat <- sleepstudy %>%
  mutate(Period = factor(period,
                      ordered = TRUE),
         Days = factor(sleepstudy$Days, ordered = TRUE)
  )

str(dat)
```

# RM ANOVA on Days

## Ensuring the Days contrasts are orthogonal

```{r orthog_days}
dat$Days %>%
  contrasts() %>%
  cor() %>% # sub cor() w/ crossprod() if you would like to check orthonormality
  round(3)
```

## Exploratory Plot (Days vs. Reaction for each Subject)

```{r plot1}
day_plot <- dat %>%
  ggplot(aes(x = Days, y = Reaction, group = 1)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "purple") +
  facet_wrap(~Subject, nrow = 3)
```

## Subjects have daily measurements

```{r day_plot}
day_plot
```

## Use aov and Error for RM ANOVA on Days

```{r aov_days}
# Base r solution
aov_days <- aov(Reaction~Days + Error(Subject/Days), 
                  data = dat)
summary(aov_days)
```

## Use ezANOVA for RM ANOVA on Days

```{r ez_days}
# ez package solution (checks for violations of sphericity)
# ----------
# This DOES NOT throw a warning because there is a single measurement for each
# Subject * Day combination
ez_days <- ezANOVA(data = dat, dv = .(Reaction),
                  within = .(Days), wid = .(Subject),
                  detailed = TRUE)
ez_days
```

# RM ANOVA on Period

## Ensuring the Period contrasts are orthogonal

```{r orthog_period}
dat$Period %>%
  contrasts() %>%
  cor() %>% # sub cor() w/ crossprod() if you would like to check orthonormality
  round(3)
```

## Exploratory Plot (Period vs. Reaction for each Subject)

```{r plot2}
period_plot <- dat %>%
  ggplot(aes(x = Period, y = Reaction, group = 1)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "purple") + 
  facet_wrap(~Subject, nrow = 3)
```

## Looking at our recoded factor, Period

```{r period_plot}
period_plot
```

## Use aov and Error for RM ANOVA on Period

```{r aov_period}
aov_period <- aov(Reaction~Period + Error(Subject/Period), 
               data = dat)
summary(aov_period)
```

## Use ezANOVA for RM ANOVA on Period

```{r ez_period, warning = TRUE}
# This throws a warning because there are multiple measurements for each
# Subject * Period combination. RM ANOVA requires single measurement for each
# combination of within-subjects factors
ez_period <- ezANOVA(data = dat, dv = .(Reaction),
                  within = .(Period), wid = .(Subject),
                  detailed = TRUE)
```

## Aggregate the data before running a traditional RM ANOVA

```{r data_grp}
dat_grp <- dat %>%
  group_by(Period, Subject) %>%
  summarize(Reaction = mean(Reaction))

# Now model the aggregate data
ez_grp <- ezANOVA(data = dat_grp, dv = .(Reaction),
                     within = .(Period), wid = .(Subject),
                     detailed = TRUE)
```

## View the ezANOVA output for aggregated period

```{r view_output}
ez_grp
```

# Looking ahead... the wonderful world of lmer

## lmer (linear mixed effects regression)

```{r lmer}
# The following model will not run because there is no covariance matrix to
# estimate in the model (i.e., no variance of days within subjects)
#lmer(Reaction ~ Days + (1|Subject/Days), data = dat)

# lmer expects unaggragated data, to model the means AND covariances of 
# parameters. Note that this is the opposite of the traditional RM ANOVA.
# Traditional RM ANOVA expects aggregated subject * condition data!
lmer_period <- lmer(Reaction ~ Period + (1|Subject/Period), data = dat)
summary(lmer_period)$coefficients
```

## Comparing lmer results to RM ANOVA results

```{r comp_res}
(ez_f <- ez_period$ANOVA %>%
  filter(Effect == "Period") %>%
  select(`F`) %>%
  pull())

(lmer_t <- summary(lmer_period)$coefficients[-1, "t value"]^2 %>%
  mean())

round(ez_f, 2) == round(lmer_t, 2)
```

# Now it's your turn!

Download the in-class activity and answer the questions as best you can. 

Collaborate with your peers to write code. 

I am available for programming questions.